---
title: "Higgs-ML summary"
author: "Rail Suleymanov"
date: "Monday, February 02, 2015"
output:
  html_document:
    keep_md: yes
---

# Load data
## Training data
Read training data, find number of missing points for each variable, remove missing points.
```{r}
train_f <- "training.csv"
train <- read.csv(train_f)
train_no_na <- data.frame(train)
features <- colnames(train)
nas1 <- vector("integer", length = length(features))
for (feature in seq_along(features)) {
    ina <- train[, feature] == -999.0
    nas1[feature] <- sum(ina)
}
for (feature in seq_along(features)) {
    ina <- train_no_na[, feature] == -999.0
    train_no_na <- train_no_na[!ina, ]
}
```

Number of missing values for each variable.
```{r}
library(ggplot2)
nas <- data.frame(features, nas1)
nas <- nas[nas1 != 0, ]
q <- qplot(x = features, y = nas1, data = nas, geom = "bar", stat = "identity")
q + theme(axis.text.x = element_text(angle = 90)) + xlab("Number of missing values")
```

## Test data
Read test data, find number of missing points for each variable, remove missing points.
```{r}
test_f <- "test.csv"
test <- read.csv(test_f)
test_no_na <- data.frame(test)
features <- colnames(test)
nas2 <- vector("integer", length = length(features))
for (feature in seq_along(features)) {
    ina <- test[, feature] == -999.0
    nas2[feature] <- sum(ina)
}
for (feature in seq_along(features)) {
    ina <- test_no_na[, feature] == -999.0
    test_no_na <- test_no_na[!ina, ]
}
```

Number of missing values for each variable.
```{r}
nas <- data.frame(features, nas2)
nas <- nas[nas2 != 0, ]
q <- qplot(x = features, y = nas2, data = nas, geom = "bar", stat = "identity")
q + theme(axis.text.x = element_text(angle = 90)) + xlab("Number of missing values")
```

# Details of some of features
```{r}
s_train <- summary(train_no_na)
s_test <- summary(test_no_na)
```

## 1) Estimated mass of the Higgs boson candidate (DER_mass_MMC)
```{r}
main <- "Histogram of DER_mass_MMC"
ylab <- "Frequency"
leg <- c("Test", "Train")
hist(test_no_na$DER_mass_MMC, breaks = 50, col = "red", ylab = ylab, main = main, xlim = c(0, 500))
hist(train_no_na$DER_mass_MMC, breaks = 50, col = "blue", add = T)
```

Test set statistics: `r s_test[, 2]`

Train set statistics: `r s_train[, 2]`

## 2) Transverse mass between the missing transverse energy and the lepton (DER_mass_transverse_met_lep)
```{r}
main <- "Histogram of DER_mass_transverse_met_lep"
ylab <- "Frequency"
leg <- c("Test", "Train")
hist(test_no_na$DER_mass_transverse_met_lep, 
     breaks = 50, col = "red", ylab = ylab, main = main, xlim = c(0, 200))
hist(train_no_na$DER_mass_transverse_met_lep, breaks = 50, col = "blue", add = T)
```

Test set statistics: `r s_test[, 3]`

Train set statistics: `r s_train[, 3]`

## 3) Invariant mass of the hadronic tau and the lepton (DER_mass_vis)
```{r}
main <- "Histogram of DER_mass_vis"
ylab <- "Frequency"
leg <- c("Test", "Train")
hist(test_no_na$DER_mass_vis, breaks = 50, col = "red", ylab = ylab, main = main, xlim = c(0, 200))
hist(train_no_na$DER_mass_vis, breaks = 50, col = "blue", add = T)
```

Test set statistics: `r s_test[, 4]`

Train set statistics: `r s_train[, 4]`

## 4) Modulus of the vector sum of the transverse momentum of the hadronic tau, the lepton,
## and the missing transverse energy vector (DER_pt_h)
```{r}
main <- "Histogram of DER_pt_h"
ylab <- "Frequency"
leg <- c("Test", "Train")
hist(test_no_na$DER_pt_h, breaks = 50, col = "red", ylab = ylab, main = main, xlim = c(0, 500))
hist(train_no_na$DER_pt_h, breaks = 50, col = "blue", add = T)
```

Test set statistics: `r s_test[, 5]`

Train set statistics: `r s_train[, 5]`

## 5) Absolute value of the pseudorapidity separation between the two jets (DER_deltaeta_jet_jet)
```{r}
main <- "Histogram of DER_deltaeta_jet_jet"
ylab <- "Frequency"
leg <- c("Test", "Train")
hist(test_no_na$DER_deltaeta_jet_jet, breaks = 50, col = "red", ylab = ylab, main = main, xlim = c(0, 10))
hist(train_no_na$DER_deltaeta_jet_jet, breaks = 50, col = "blue", add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Test set statistics: `r s_test[, 6]`

Train set statistics: `r s_train[, 6]`
