---
title: "Higgs-ML summary"
author: "Rail Suleymanov"
date: "Monday, February 02, 2015"
output:
  html_document:
    keep_md: yes
---

# Load data
## Training data
Read training data, find number of missing points for each variable, remove missing points.
```{r}
train_f <- "training.csv"
train <- read.csv(train_f)
train_no_na <- data.frame(train)
features <- colnames(train)
nas1 <- vector("integer", length = length(features))
for (feature in seq_along(features)) {
    ina <- train[, feature] == -999.0
    nas1[feature] <- sum(ina)
}
for (feature in seq_along(features)) {
    ina <- train_no_na[, feature] == -999.0
    train_no_na <- train_no_na[!ina, ]
}
```

Number of missing values for each variable.
```{r}
library(ggplot2)
nas <- data.frame(features, nas1)
nas <- nas[nas1 != 0, ]
q <- qplot(x = features, y = nas1, data = nas, geom = "bar", stat = "identity")
q + theme(axis.text.x = element_text(angle = 90)) + xlab("Number of missing values")
```

Check if there are full matching na's in different features. That can be the case in variables **DER_deltaeta_jet_jet** (6-th variable), **DER_mass_jet_jet** (7-th), **DER_prodeta_jet_jet** (8-th), **DER_lep_eta_centrality** (14-th), **PRI_jet_subleading_pt** (28-th), **PRI_jet_subleading_eta** (29-th), **PRI_jet_subleading_phi** (30-th) and in variables **PRI_jet_leading_pt** (25-th), **PRI_jet_leading_eta** (26-th), **PRI_jet_leading_phi** (27-th).
```{r}
ina6 <- which(train$DER_deltaeta_jet_jet == -999)
ina7 <- which(train$DER_mass_jet_jet == -999)
ina8 <- which(train$DER_prodeta_jet_jet == -999)
ina14 <- which(train$DER_lep_eta_centrality == -999)
ina28 <- which(train$PRI_jet_subleading_pt == -999)
ina29 <- which(train$PRI_jet_subleading_eta == -999)
ina30 <- which(train$PRI_jet_subleading_phi == -999)

match_res <- sum(intersect(ina6, ina7) == ina6) == length(ina7)
match_res <- sum(intersect(ina6, ina8) == ina6) == length(ina8)
match_res <- sum(intersect(ina6, ina14) == ina6) == length(ina14)
match_res <- sum(intersect(ina6, ina28) == ina6) == length(ina28)
match_res <- sum(intersect(ina6, ina29) == ina6) == length(ina29)
match_res <- sum(intersect(ina6, ina30) == ina6) == length(ina30)
```

Do all na's in those variables match? `r match_res`

```{r}
ina25 <- which(train$PRI_jet_leading_pt == -999.0)
ina26 <- which(train$PRI_jet_leading_eta == -999.0)
ina27 <- which(train$PRI_jet_leading_phi == -999.0)

match_res <- sum(intersect(ina25, ina26) == ina25) == length(ina26)
match_res <- sum(intersect(ina25, ina27) == ina27) == length(ina27)
```

Do all na's in those variables match? `r match_res`

We see that there's indeed full matching of these features. Now let's explore how do indexes of NA features intersect for **DER_mass_MMC**, **DER_deltaeta_jet_jet** and **PRI_jet_leading_pt**.
```{r}
library(VennDiagram)
ina2 <- which(train$DER_mass_MMC == -999)
area1 <- length(ina6)
area2 <- length(ina2)
area3 <- length(ina25)
in12 <- length(intersect(ina6, ina2))
in13 <- length(intersect(ina6, ina25))
in23 <- length(intersect(ina2, ina25))
in123 <- length(intersect(ina6, intersect(ina2, ina25)))

venn.plot <- draw.triple.venn(area1, area2, area3, in12, in23, in13, in123, 
                              category = c("DER_deltaeta_jet_jet", "DER_mass_MMC", "PRI_jet_leading_pt"),
                              fill = c("blue", "red", "green"),
                              lty = "blank", cat.col = rep("black", 3))
```


## Test data
Read test data, find number of missing points for each variable, remove missing points.
```{r}
test_f <- "test.csv"
test <- read.csv(test_f)
test_no_na <- data.frame(test)
features <- colnames(test)
nas2 <- vector("integer", length = length(features))
for (feature in seq_along(features)) {
    ina <- test[, feature] == -999.0
    nas2[feature] <- sum(ina)
}
for (feature in seq_along(features)) {
    ina <- test_no_na[, feature] == -999.0
    test_no_na <- test_no_na[!ina, ]
}
```

Number of missing values for each variable.
```{r}
nas <- data.frame(features, nas2)
nas <- nas[nas2 != 0, ]
q <- qplot(x = features, y = nas2, data = nas, geom = "bar", stat = "identity")
q + theme(axis.text.x = element_text(angle = 90)) + xlab("Number of missing values")
```

Perform same na's matching for test set as it was for training set
```{r}
ina6 <- which(test$DER_deltaeta_jet_jet == -999)
ina7 <- which(test$DER_mass_jet_jet == -999)
ina8 <- which(test$DER_prodeta_jet_jet == -999)
ina14 <- which(test$DER_lep_eta_centrality == -999)
ina28 <- which(test$PRI_jet_subleading_pt == -999)
ina29 <- which(test$PRI_jet_subleading_eta == -999)
ina30 <- which(test$PRI_jet_subleading_phi == -999)

match_res <- sum(intersect(ina6, ina7) == ina6) == length(ina7)
match_res <- sum(intersect(ina6, ina8) == ina6) == length(ina8)
match_res <- sum(intersect(ina6, ina14) == ina6) == length(ina14)
match_res <- sum(intersect(ina6, ina28) == ina6) == length(ina28)
match_res <- sum(intersect(ina6, ina29) == ina6) == length(ina29)
match_res <- sum(intersect(ina6, ina30) == ina6) == length(ina30)
```

Do all na's in those variables match? `r match_res`

```{r}
ina25 <- which(test$PRI_jet_leading_pt == -999.0)
ina26 <- which(test$PRI_jet_leading_eta == -999.0)
ina27 <- which(test$PRI_jet_leading_phi == -999.0)

match_res <- sum(intersect(ina25, ina26) == ina25) == length(ina26)
match_res <- sum(intersect(ina25, ina27) == ina27) == length(ina27)
```

Do all na's in those variables match? `r match_res`

We see that there's indeed full matching of these features. Now let's explore how do indexes of NA features intersect for **DER_mass_MMC**, **DER_deltaeta_jet_jet** and **PRI_jet_leading_pt**.
```{r}
library(VennDiagram)
ina2 <- which(test$DER_mass_MMC == -999)
area1 <- length(ina6)
area2 <- length(ina2)
area3 <- length(ina25)
in12 <- length(intersect(ina6, ina2))
in13 <- length(intersect(ina6, ina25))
in23 <- length(intersect(ina2, ina25))
in123 <- length(intersect(ina6, intersect(ina2, ina25)))

venn.plot <- draw.triple.venn(area1, area2, area3, in12, in23, in13, in123, 
                              category = c("DER_deltaeta_jet_jet", "DER_mass_MMC", "PRI_jet_leading_pt"),
                              fill = c("blue", "red", "green"),
                              lty = "blank", cat.col = rep("black", 3))
```

# Conclusion
According to NA's structure investigation, 2 strategies are possible:

1. As there's complete NA-matching between several sets of variables, we can conclude that whole dataset can be divided to following parts

 - examples having all variables defined (i.e. without indexes that go into coloured regions on the last figure) - set A

 - examples that didn't go to 'set A', having all variables defined except variable **DER_mass_MMC** (coloured red) - set B

 - examples that didn't go to previous sets, having all variables defined except variable **DER_deltaeta_jet_jet** (coloured blue) - set C
 
 - examples that didn't go to previous sets, having all variables defined except variables **DER_mass_MMC** AND **DER_deltaeta_jet_jet** (coloured purple) - set D

 - examples that didn't go to previous sets, having all variables defined except variables **DER_deltaeta_jet_jet** AND **PRI_jet_leading_pt** (light green) - set E

 - examples that didn't go to previous sets, having all variables defined except variables **DER_mass_MMC**, **DER_deltaeta_jet_jet** AND **PRI_jet_leading_pt** (dark green, the intersection of all 3 regions) - set F
 
2. Leave dataset 'asis' but then need to fill NA's using algorithm one of

 - variable means
 
 - generate randoms from that variable's distributions
 
 - k-nearest neighbors
 
 - k-means clustering

# Details of some of features
We will choose subset from train data (without NAs) and plot its details.
```{r}
s_train <- summary(train_no_na)
ind <- sample(1:dim(train_no_na)[1], size = 5000, replace = F)
sub <- train_no_na[ind, ]
s_inds <- sub$Label == "s"
b_inds <- sub$Label == "b"
num_s <- length(sub$Label[s_inds])
num_b <- length(sub$Label[b_inds])
```

## 1) Estimated mass of the Higgs boson candidate (DER_mass_MMC)
```{r}
main <- "Histogram of DER_mass_MMC"
xlab <- "DER_mass_MMC"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_mass_MMC[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 500))
hist(sub$DER_mass_MMC[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_mass_MMC[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_mass_MMC[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 2]`

## 2) Transverse mass between the missing transverse energy and the lepton (DER_mass_transverse_met_lep)
```{r}
main <- "Histogram of DER_mass_transverse_met_lep"
xlab <- "DER_mass_transverse_met_lep"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_mass_transverse_met_lep[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 250))
hist(sub$DER_mass_transverse_met_lep[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_mass_transverse_met_lep[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_mass_transverse_met_lep[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 3]`

## 3) Invariant mass of the hadronic tau and the lepton (DER_mass_vis)
```{r}
main <- "Histogram of DER_mass_vis"
xlab <- "DER_mass_vis"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_mass_vis[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 250))
hist(sub$DER_mass_vis[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_mass_vis[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_mass_vis[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 4]`

## 4) Modulus of the vector sum of the transverse momentum of the hadronic tau, the lepton,
## and the missing transverse energy vector (DER_pt_h)
```{r}
main <- "Histogram of DER_pt_h"
xlab <- "DER_pt_h"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_pt_h[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 500))
hist(sub$DER_pt_h[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_pt_h[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_pt_h[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 5]`

## 5) Absolute value of the pseudorapidity separation between the two jets (DER_deltaeta_jet_jet)
```{r}
main <- "Histogram of DER_deltaeta_jet_jet"
xlab <- "DER_deltaeta_jet_jet"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_deltaeta_jet_jet[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 10))
hist(sub$DER_deltaeta_jet_jet[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_deltaeta_jet_jet[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_deltaeta_jet_jet[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 6]`

## 6) The invariant mass of the two jets (DER_mass_jet_jet)
```{r}
main <- "Histogram of DER_mass_jet_jet"
xlab <- "DER_mass_jet_jet"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_mass_jet_jet[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 3000))
hist(sub$DER_mass_jet_jet[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_mass_jet_jet[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_mass_jet_jet[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 7]`

## 7) The product of the pseudorapidities of the two jets (DER_prodeta_jet_jet)
```{r}
main <- "Histogram of DER_prodeta_jet_jet"
xlab <- "DER_prodeta_jet_jet"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_prodeta_jet_jet[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-15, 15))
hist(sub$DER_prodeta_jet_jet[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_prodeta_jet_jet[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_prodeta_jet_jet[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 8]`

## 8) The R separation between the hadronic tau and the lepton (DER_deltar_tau_lep)
```{r}
main <- "Histogram of DER_deltar_tau_lep"
xlab <- "DER_deltar_tau_lep"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_deltar_tau_lep[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 4))
hist(sub$DER_deltar_tau_lep[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_deltar_tau_lep[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_deltar_tau_lep[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 9]`

## 9) The modulus of the vector sum of the missing transverse momenta and the transverse momenta of th ehadronic tau, the lepton, the leading jet and the subleading jet (DER_pt_tot)
```{r}
main <- "Histogram of DER_pt_tot"
xlab <- "DER_pt_tot"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_pt_tot[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 200))
hist(sub$DER_pt_tot[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_pt_tot[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_pt_tot[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 10]`

## 10) The sum of the moduli of the transverse momenta of the hadronic tau, the lepton, the leading jet and the subleading jet (DER_sum_pt)
```{r}
main <- "Histogram of DER_sum_pt"
xlab <- "DER_sum_pt"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_sum_pt[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 1000))
hist(sub$DER_sum_pt[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_sum_pt[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_sum_pt[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 11]`

## 11) The ratio of the transverse momenta of the lepton and the hadronic tau (DER_pt_ratio_lep_tau)
```{r}
main <- "Histogram of DER_pt_ratio_lep_tau"
xlab <- "DER_pt_ratio_lep_tau"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_pt_ratio_lep_tau[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 10))
hist(sub$DER_pt_ratio_lep_tau[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_pt_ratio_lep_tau[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_pt_ratio_lep_tau[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 12]`

## 12) The centrality of the azimutal angle of the missing transverse energy vector wrt the hadronic tau and the lepton (DER_met_phi_centrality)
```{r}
main <- "Histogram of DER_met_phi_centrality"
xlab <- "DER_met_phi_centrality"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_met_phi_centrality[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-1.5, 1.5))
hist(sub$DER_met_phi_centrality[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_met_phi_centrality[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_met_phi_centrality[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 13]`

## 13) The centrality of the pseudorapidity of the lepton wrt the two jets (DER_lep_eta_centrality)
```{r}
main <- "Histogram of DER_lep_eta_centrality"
xlab <- "DER_lep_eta_centrality"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$DER_lep_eta_centrality[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 1))
hist(sub$DER_lep_eta_centrality[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$DER_lep_eta_centrality[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$DER_lep_eta_centrality[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 14]`

## 14) The transverse momentum of the hadronic tau (PRI_tau_pt)
```{r}
main <- "Histogram of PRI_tau_pt"
xlab <- "PRI_tau_pt"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_tau_pt[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 300))
hist(sub$PRI_tau_pt[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_tau_pt[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_tau_pt[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 15]`

## 15) The pseudorapidity of the hadronic tau (PRI_tau_eta)
```{r}
main <- "Histogram of PRI_tau_eta"
xlab <- "PRI_tau_eta"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_tau_eta[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-3, 3))
hist(sub$PRI_tau_eta[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_tau_eta[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_tau_eta[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 16]`

## 16) The azimuth angle of the hadronic tau (PRI_tau_phi)
```{r}
main <- "Histogram of PRI_tau_phi"
xlab <- "PRI_tau_phi"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_tau_phi[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-3, 3))
hist(sub$PRI_tau_phi[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_tau_phi[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_tau_phi[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 17]`

## 17) The transverse momentum of the lepton - electron or muon (PRI_lep_pt)
```{r}
main <- "Histogram of PRI_lep_pt"
xlab <- "PRI_lep_pt"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_lep_pt[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 250))
hist(sub$PRI_lep_pt[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_lep_pt[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_lep_pt[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 18]`

## 18) The pseudorapidity of the lepton (PRI_lep_eta)
```{r}
main <- "Histogram of PRI_lep_eta"
xlab <- "PRI_lep_eta"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_lep_eta[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-3, 3))
hist(sub$PRI_lep_eta[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_lep_eta[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_lep_eta[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 19]`

## 19) The azimuth angle of the lepton (PRI_lep_phi)
```{r}
main <- "Histogram of PRI_lep_phi"
xlab <- "PRI_lep_phi"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_lep_phi[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-3, 3))
hist(sub$PRI_lep_phi[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_lep_phi[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_lep_phi[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 20]`

## 20) The missing transverse energy (PRI_met)
```{r}
main <- "Histogram of PRI_met"
xlab <- "PRI_met"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_met[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 400))
hist(sub$PRI_met[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_met[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_met[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 21]`

## 21) The azimuth angle of the missing transverse energy (PRI_met_phi)
```{r}
main <- "Histogram of PRI_met_phi"
xlab <- "PRI_met_phi"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_met_phi[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-3, 3))
hist(sub$PRI_met_phi[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_met_phi[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_met_phi[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 22]`

## 22) The total transverse energy in the detector (PRI_met_sumet)
```{r}
main <- "Histogram of PRI_met_sumet"
xlab <- "PRI_met_sumet"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_met_sumet[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 1000))
hist(sub$PRI_met_sumet[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_met_sumet[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_met_sumet[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 23]`

## 23) The transverse momentum of the leading jet, that is the jet with largest transverse momentum (PRI_jet_leading_pt)
```{r}
main <- "Histogram of PRI_jet_leading_pt"
xlab <- "PRI_jet_leading_pt"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_jet_leading_pt[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 600))
hist(sub$PRI_jet_leading_pt[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_jet_leading_pt[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_jet_leading_pt[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 24]`

## 24) The pseudorapidity of the leading jet (PRI_jet_leading_eta)
```{r}
main <- "Histogram of PRI_jet_leading_eta"
xlab <- "PRI_jet_leading_eta"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_jet_leading_eta[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-4, 4))
hist(sub$PRI_jet_leading_eta[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_jet_leading_eta[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_jet_leading_eta[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 25]`

## 25) The azimuth angle of the leading jet (PRI_jet_leading_phi)
```{r}
main <- "Histogram of PRI_jet_leading_phi"
xlab <- "PRI_jet_leading_phi"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_jet_leading_phi[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-3, 3))
hist(sub$PRI_jet_leading_phi[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_jet_leading_phi[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_jet_leading_phi[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 26]`

## 26) The transverse momentum of the leading jet, that is, the jet with second largest transverse momentum (PRI_jet_subleading_pt)
```{r}
main <- "Histogram of PRI_jet_subleading_pt"
xlab <- "PRI_jet_subleading_pt"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_jet_subleading_pt[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 250))
hist(sub$PRI_jet_subleading_pt[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_jet_subleading_pt[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_jet_subleading_pt[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 27]`

## 27) The pseudorapidity of the subleading jet (PRI_jet_subleading_eta)
```{r}
main <- "Histogram of PRI_jet_subleading_eta"
xlab <- "PRI_jet_subleading_eta"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_jet_subleading_eta[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-5, 5))
hist(sub$PRI_jet_subleading_eta[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_jet_subleading_eta[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_jet_subleading_eta[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 28]`

## 28) The azimuth angle of the subleading jet (PRI_jet_subleading_phi)
```{r}
main <- "Histogram of PRI_jet_subleading_phi"
xlab <- "PRI_jet_subleading_phi"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_jet_subleading_phi[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(-3, 3))
hist(sub$PRI_jet_subleading_phi[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_jet_subleading_phi[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_jet_subleading_phi[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 29]`

## 29) The scalar sum of the transverse momentum of all the jets of the events (PRI_jet_all_pt)
```{r}
main <- "Histogram of PRI_jet_all_pt"
xlab <- "PRI_jet_all_pt"
ylab <- "Frequency"
leg <- c("Signal", "Background")
par(mfrow = c(1, 2))
hist(sub$PRI_jet_all_pt[s_inds], breaks = 50, col = rgb(1, 0, 0, 0.5), 
     xlab = xlab, ylab = ylab, main = main, xlim = c(0, 800))
hist(sub$PRI_jet_all_pt[b_inds], breaks = 50, col = rgb(0, 0, 1, 0.5), add = T)
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
plot(1:num_s, sub$PRI_jet_all_pt[s_inds], col = "red", xlim = c(0, 5000))
points((num_s+1):5000, sub$PRI_jet_all_pt[b_inds], col = "blue")
legend("topright", pch = 1, col = c("red", "blue"), legend = leg)
```

Train set statistics: `r s_train[, 29]`
